### 2021-12-07 doc/contrastive-learning.tex
adds first version of proposal

\section{Addressing Limitations}

\subsection{Supervised Contrastive Learning}

@@

\subsection{Clustering}

@@

\subsection{Memory Banks \& Momentum Encoders}

@@

\subsection{Hard Negative \& Positive Mining}

@@

\begin{figure}[h!]
    \begin{center}
        \includegraphics[scale=0.5]{images/synthetic-negatives.png}
        \caption{Negative Hard Mixing \cite{kalantidis_hard_nodate}}
    \end{center}
\end{figure}

@@

>>>>

\section{Addressing Limitations}
\label{limitations}

\subsection{Supervised Contrastive Learning}
\label{limitations:supervised}

@@

\subsection{Clustering}
\label{limitations:clustering}

@@

\subsection{Memory Banks \& Momentum Encoders}
\label{limitations:memory}

@@

\subsection{Hard Negative \& Positive Mining}
\label{limitations:mining}

@@

\begin{figure}[h!]
    \begin{center}
        \includegraphics[scale=0.35]{images/synthetic-negatives.png}
        \caption{Negative Hard Mixing \cite{kalantidis_hard_nodate}}
    \end{center}
\end{figure}

@@

My work is focused on improving self-supervised contrastive learning in 3D point clouds for semantic and instance segmentation. I will use the work of PointContrast as my starting point. My focus will be on improving the Positive and Negative sampling strategies so that the pretext task and downstream performance improves as well as requiring fewer computational resources.

It's been shown that InfoNCE loss with a temperature parameter is the most effective contrastive loss and the main one used in all recent vision applications. For this reason I will stick with this loss function. It is reprinted below for completeness.

\subsection{Negative Sampling}

The main focus will be on improving the negative sampling techniques. Currently, negative samples are selected uniformly at random from the set of positive keys. However, this is a naive sampling strategy that is not conditioned on the sampled query point at all. It introduces the possibility for a large number of false negatives or semi-false negatives. Ideally we want to sample from a distribution like:

$$\mathbf{x}^{-} \sim p^{-}( \cdot | \mathbf{x})$$

To condition on the input point, I plan to use Fast Histogram Point Features (FHPF) as a way to select negative points that are sufficiently different from the input point. This operates on the assumption that if the point is different locally, then it for sure is different globally even if the surrounding area is the same.

$$p^{-}( \cdot | \mathbf{x}) = \text{Uniform}(\{ \: \mathbf{x}^- \: | \: \text{sim}(\text{FPFH}(\mathbf{x}^-),\text{FPFH}(\mathbf{x})) > \delta \})$$

where $\text{sim}(\cdot,\cdot)$ is some kind of distance function between the histograms such as earth movers distance.

\begin{figure}[h!]
    \begin{center}
        \includegraphics[scale=0.3]{images/regions.png}
        \caption{Visualization of local vs global region dissimilarity}
    \end{center}
\end{figure}

Instead of uniform sampling, I can also try cluster sampling by clustering different local features into feature types (e.g. flat surfaces, round objects, corners) and sample equally from those distributions.

$$p^{-}( \cdot | \mathbf{x}) = \text{Uniform}(\{ \textbf{c} | \textbf{c} \in \text{KNN}\{ \text{FPFH} ( \{ \textbf{x}^- | \textbf{x}^- \in \mathcal{X} \} ) \}, \textbf{c} \neq \textbf{x} \})$$

Where $\text{KNN}$ represents the K-nearest-neigbours clustering algorithm and $\textbf{c}$ is a set of negative samples clustered together based on their similar properties. Thus this strategy constitutes sampling equally from different clusters of points.
%

Another option is to weight the negative samples based off of the local dissimilarity of the points and incorporate this into the loss function itself.
%

$$\mathcal{L}_{i} = -\log \frac{\exp(q_{i} \cdot k^{+}_{i} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + \sum^{K}_{j}\exp(w_{i,j} (q_{i} \cdot k^{-}_{j}) / \tau)}$$

$$\text{where} \: w_{i,j} = 1 - \frac{s_{i,j}}{\sum_k^{K}s_{i,k}} $$
$$ s_{i,j} = \text{sim}(\text{FPFH}(q_{i}), \text{FPFH}(k^{-}_{j}))$$

\begin{figure}[h!]
    \begin{center}
        \includegraphics[scale=0.5]{images/fpfh.png}
        \caption{Visualization of Fast Point Feature Histogram features in a sample point cloud}
    \end{center}
\end{figure}

The last possibility is to look at applying different loss functions at different parts of the network depending on the receptive field. So that the parts looking at local regions have different objects than those looking at more global regions. Perhaps there's no advantage to training the global region at all?

\subsection{Positive Sampling}

Using different augmentations or adding points from the same cluster as the query.

\subsection{Informative Sampling}

\subsection{Other Techniques}

Implement all the negative sampling techniques from section \ref{limitations:mining} as well as compare using a momentum encoder and perhaps one of the clustering techniques from section \ref{limitations:clustering}
